---
title: "task3_data_analysis"
author: "Dan McG"
date: "21/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will be carrying out some analysis on sea bird observations

First lets load our libriaries
```{r}
library(tidyverse)
library(here)
```

Now lets load the data set
```{r}
seabird_df <- read_csv(here("clean_data/seabird_data_clean.csv"))
```

Which bird had the most individual sightings?

```{r}
seabird_df %>%
  group_by(species_common_name) %>%
  summarise(individual_sightings = n()) %>%
  arrange(desc(individual_sightings)) %>%
  head(1)
```

The Wandering Alabtross sensi lato PL2 was had the most individual sightings with 3104 different observations

Which bird had the highest total count?

```{r}
seabird_df %>%
  group_by(species_common_name) %>%
  summarise(total_count = sum(COUNT)) %>%
  arrange(desc(total_count)) %>%
  head(1)
```

The Black-billed gull had the hightest total count with 7732 observed

Which bird had the highest total count above a latitude of -30?
```{r}
seabird_df %>%
  filter(LAT > -30) %>%
  group_by(species_common_name) %>%
  summarise(total_count = sum(COUNT)) %>%
  arrange(desc(total_count)) %>%
  head(1)
```

The Wedge-tailed shearwater has the highest total count above a latitude of -30 with 855 observed in total

How many different types of birds were only ever seen in groups of 1?
```{r}
single_sightings <- seabird_df %>%
  group_by(species_common_name) %>% 
  summarise(total_count = sum(COUNT), individual_sightings = n()) %>% # make additional columns that give the total number of birds oberved and the number of individual sightings
  filter(total_count/individual_sightings == 1) # if the total number of birds observed is the same as the number of individual sightings then these birds must have only been observed by themselves

count_lonely_birds <- length(unique(single_sightings$species_common_name))

count_lonely_birds    
```

73 different types of birds have only ever been spotted in groups of 1

How many penguins were seen?
```{r}
penguin_df <- seabird_df[grep("[::punct:: ]*[Pp]enguin[ ::punct::]*",seabird_df$species_common_name, ), ] # make a separate df for all species of penguins

no_penguin_species <- length(unique(penguin_df$species_common_name))

total_no_oberved_penguins <- sum(penguin_df$COUNT, na.rm = TRUE)


no_penguin_species
total_no_oberved_penguins
```

There are 9 individual penguin species recorded including unidentified penguin sightings

There has been a total of 158 penguins recorded



