---
title: "R Notebook"
output: html_notebook
---

Import our libriaries
```{r}
library(tidyverse)
library(here)
```

Import Data set
```{r}
halloween_df <- read_csv(here("clean_data/halloween_candy_clean.csv"), col_types = c(col_character(),
                                                                                    col_character(),
                                                                                    col_double(),
                                                                                    col_character(),
                                                                                    col_character(),
                                                                                    col_character(),
                                                                                    col_character(),
                                                                                    col_character(),
                                                                                    col_character(),
                                                                                    col_character(),
                                                                                    col_double()
                                                                                    ))
```

1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
```{r}
halloween_df %>%
  summarise(total_candy_ratings = sum(!is.na(Q06_response)))
```

511242 total candy ratings

2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
halloween_df %>%
  group_by(Q01_going_out) %>%
  summarise(average_age = round(mean(Q03_age, na.rm = TRUE)))
```

The average age of those not going out is 39 and the average age of those going out is 35

3. For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r}
response_count <- halloween_df %>% 
  group_by(Q06_candy_name, Q06_response) %>%
  count(Q06_response, sort = TRUE) %>%
  top_n(n = 1)

# filter for each response
response_count %>%
  filter(Q06_response == "JOY") %>% head(1) # JOY

response_count %>%
  filter(Q06_response == "DESPAIR") %>% head(1) # DESPAIR

response_count %>%
  filter(Q06_response == "MEH") %>% head(1) # MEH
```

Lollipops are the most meh out of the candy's
Mary Janes bring the most despair
Reese's Peanut Butter Cups brings the most joy


For the next three questions, count despair as -1, joy as +1 and meh as 0.
```{r}
rated_halloweenn_df <- halloween_df %>%
  mutate(Q06_response =  gsub("JOY", 1, Q06_response)) %>%
   mutate(Q06_response =  gsub("DESPAIR", -1, Q06_response)) %>%
   mutate(Q06_response =  gsub("MEH", 0, Q06_response))

rated_halloweenn_df$Q06_response <- as.numeric(rated_halloweenn_df$Q06_response)

```


5. What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
rated_halloweenn_df %>%
  group_by(Q02_gender, Q06_candy_name) %>%
  summarise(Total = sum(Q06_response, na.rm = TRUE)) %>%
  top_n(n = 1)
```

6. What was the most popular candy bar in each year?
```{r}
rated_halloweenn_df %>%
  group_by(Q10_year, Q06_candy_name) %>%
  summarise(Total = sum(Q06_response, na.rm = TRUE)) %>%
  top_n(n = 1)
```


7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
rated_halloweenn_df %>%
  filter(Q04_country == "America" | Q04_country == "UK" | Q04_country == "Canada" ) %>%
  group_by(Q04_country, Q06_candy_name) %>%
  summarise(Total = sum(Q06_response, na.rm = TRUE)) %>%
  top_n(n = 1)

# check for rest of world
rated_halloweenn_df %>%
  filter(Q04_country != "America" & Q04_country != "UK" & Q04_country != "Canada" ) %>%
   group_by(Q06_candy_name) %>%
  summarise(row_Total = sum(Q06_response, na.rm = TRUE)) %>%
  top_n(n = 1)
```

