---
title: "R Notebook"
output: html_notebook
---

Load libriaries
```{r}
library(tidyverse)
library(here)
```

Load Data Set
```{r}
dec_data_clean <- read_csv(here("clean_data/decathlon_data_clean.csv"))
```

1. Who had the longest long jump seen in the data?
```{r}
dec_data_clean %>%
  arrange(desc(long_jump)) %>% # arrabge rows by long jump distance descending
  head(1) # limit the results to 1 to give the best performer in the data set
```

Clay had the longest long jump

2. What was the average 100m time in each competition?

```{r}
dec_data_clean %>%
  group_by(competition) %>% # group by the different competitions the the data set
  summarise(average_100m = round(mean(run_100m), 2)) # calculate the mean for each category and format the same as the data
```

3. Who had the highest total points across both competitions?

```{r}
dec_data_clean %>%
  arrange(desc(points)) %>% # arrabge rows by event points descending
  head(1) # limit the results to 1 to give the best performer in the data set
```

Serbie had the highest points across both competitions

4. What was the shot-put scores for the top three competitors in each competition?
```{r}
dec_data_clean %>%
  arrange(rank) %>% # arange by rank
  select(athlete_name, shot_put, rank, competition) %>% # select relevant columns
  head(6) # limit to the top 6 results in order to get the data for the top 3 competitors in each competition
```

5. What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
dec_data_clean %>%
  mutate(over_50_in_400m = (`400m` > 50)) %>%
  group_by(over_50_in_400m) %>% 
  summarise(average_points = mean(points))
```

